---
title: "p8105_hw3_ym3077"
output: github_document
---

Load packages
```{r}
library(tidyverse)
```

## Problem 1

```{r}
library(p8105.datasets)
data("instacart")

instacart_summary = list(instacart)
instacart_summary
```
This `instacart` dataset has `r nrow(instacart)` observations for 131,209 users and `r ncol(instacart)` variables/rows, where each row represents a product from a single order and is associated with a customer.    
The 15 variables in this dataset are: `order_id` (order identifier), `product_id` (product identifier), `add_to_cart_order` (order of each product was added to cart), `reordered` (1 if ordered before, 0 otherwise), `user_id` (customer identifier), `eval_set` (which evaluation this order belongs to), `order_number` (the order number for this user, 1=first, n=nth), `order_dow` (the day of the week when this order was placed), `order_hour_of_day` (the hour of the day when this order was placed), `days_since_prior_order` (days since the last order, capped at 30, NA if order_number=1), `product_name` (name of the product), `aisle_id`(aisle identifier), `department_id` (department identifier), `aisle` (the name of the aisle), `department` (department this product belongs to).

```{r}


```

















